---
title: "Class12"
author: "Jae Kim"
format: pdf
editor: visual
---

# Section 1. Proportion of G/G in a population

```{r}
mxl <- read.csv("MXL.csv", row.names = 1)
mxl
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl)*100
```

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes

```{r}
url <- "https://bioboot.github.io/bimm143_S23/class-material/rs8067378_ENSG00000172057.6.txt"

expr <- read.table(url)

nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
result <- aggregate(expr$exp ~ expr$geno, data = expr, FUN = median)
result
```

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2)
```

Making a boxplot
```{r}
x <- ggplot(expr) +
  aes(geno, exp, fill=geno) +
  geom_boxplot(notch=TRUE)
x
```

```{r}
boxplot_data <- ggplot_build(x)$data[[1]]
boxplot_data$middle
```

From the boxplot, we can see that the expression of ORMDL3 is higher if the genotype is A/A than if it is G/G. It seems like SNP does affect the expression of ORMDL3.












