---
title: "Class10"
author: "Jae Kim"
format: pdf
---

# Comparative structure analysis of ADK

> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

msa

> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

bio3d-view

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?
## Search and Retrieve ADK Structures

True

## Search and Retrieve ADK structures

```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```
> Q13. How many amino acids are in this sequence, i.e. how long is this sequence?

214

We can run BLAST with this sequence:
```{r}
# Blast or hmmer search
# b <- blast.pdb(aa)
```

```{r}
# hits <- plot(b)
```

Let's see what is in our `hits`object.

```{r}
# hits$pdb.id

hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

Now we can download all of these PDB structures:

```{r}
# Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)

```

Now I want to align and superpose these structures which are all over the place.

```{r}
pdbs <- pdbaln(files, fit=TRUE, exefile="msa")
```

Let's have a look at our `pdbs` object.
```{r}
pdbs
```

Now we have our aligned and supersposed structures we can perform all sorts of analysis on them. Let's do PCA...

```{r}
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenylate kinase X-ray structures. Each dot represents one PDB structure.

We can cluster the structures by RMSD (or any other method).

```{r}
rd <- rmsd(pdbs)

hc.rd <- hclust(dist(rd))
grps <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col=grps)
```

# Optional further visualization

We can make a wee movie - also called a trajectory of the major differences (i.e. structural displacements) of ADK.

```{r}
# Visualizae first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

